High-resolution ALMA observations of V4046\,Sgr: a circumbinary disc with a thin ring

### 2020-10-29 Draft_V4046Sgr_mnras.tex
Updates from Overleaf

% Abstract of the paper
\begin{abstract}
  The nearby V4046\,Sgr spectroscopic binary hosts a gas-rich disc known for its wide cavity and dusty ring. We present new high resolution ($\sim$20\,mas) ALMA observations of the 1.3\,mm continuum. The comparison of these observations, combined with SPHERE-IRDIS polarized images and a well-sampled spectral energy distribution (SED), against radiative transfer (RT) predictions carried out with the \textsc{radmc3d} package, allow us to propose a physical model for the source. The new ALMA data reveal a very thin ring at a radius of 13.46$\pm$0.43\,au (Ring13), with a marginally resolved radial width of 3.86$\pm$1.00\,au. Ring13 is surrounded by a $\sim$9\,au-wide gap, and it is flanked by a very bright mm outer ring (Ring24) with a sharp inner edge at 24\,au. The steeply decreasing brightness of Ring24 breaks at $\sim$35\,au into a shallow tail. The RT model requires an inner ring at $\sim$6\,au (Ring6) in small dust grains, hiding under the IRDIS coronagraph, and that surrounds an inner circumbinary disc. Faint mm-continuum close ($\sim0\farcs02$) to the stellar position is picked up by ALMA. ***
  %coincident with Ring6 is picked up by ALMA, whose morphology suggests that Ring6 is lopsided or shadowed by the secondary. 
  The previously reported scattered-light shadow of the secondary star is also reproduced by the RT model. The surprising thin Ring13 is nonetheless roughly 12 times wider than the model scale height, and could thus be long-lived. The strong near-far disc asymmetry at 1.65\,$\micron$ points at a very forward-scattering phase function, and requires grain radii of no less than 0.4\,$\micron$. 
\end{abstract}

@@

Recent observations of young circumstellar discs have transformed current knowledge of planet formation, but the focus of resolved imaging with the Atacama Large Millimetre/submillimetre Array (ALMA) or with 
%VLT/SPHERE 
the current generation of high-contrast images 
has mainly been towards the brighter sources \citep{Andrews2020arXiv200105007A}. The discs around T Tauri Stars (DARTTS) program was first presented in \citet{Avenhaus_2018} with the aim to image eight stars with the Spectro-Polarimeter High-contrast Exoplanet REsearch (SPHERE) \citep{2019A&A...631A.155B}. The sample is not biased towards exceptionally bright and large discs, and consists of only solar-type stars. A second part of the survey increased the number of sources by presenting 21 new images of circumstellar discs \citep{Garufi2020}. The observations revealed diverse structures and morphologies in the scattering surface of these discs. This letter on V4046 Sagittarii (Sgr) is part of the DARTTS survey with ALMA (DARTTS-A) which will present millimetre observations of nine protoplanetary discs previously imaged in polarized scattered light in DARTTS-S.

V4046\,Sgr is a double-lined spectroscopic binary of K-type stars (K5 and K7) with very similar masses of 0.90$\pm$0.05\,M$_{\sun}$ and 0.85$\pm$0.04\,M$_{\sun}$ \citep{Rosenfeld_2012} in a close ($a \approx 0.041$\,au), circular ($e\leq0.01$) orbit, with an orbital period of 2.42 days \citep{refId0}. It is a member of the $\beta$ Pictoris moving group \citep{Zuckerman_2004}, with an estimated age of 23$\pm$3\,Myr \citep{Mamajek_2014}, and its distance is 72.41$\pm$0.34\,pc \citep{Gaia}. V4046\,Sgr hosts a massive ($\sim$0.1\,M$_{\sun}$) circumbinary disc extending to $\sim$300\,au \citep{Rosenfeld_2013, Rodriguez_2010}, with a rich observable chemical diversity \citep{Kastner_2018}.

@@

V4046 Sgr was observed in differential polarization imaging (DPI) mode with SPHERE/IRDIS on March 13, 2016. All the details of this observation can be found in the first DARTTS-S paper \citep{Avenhaus_2018} where the image was first published. Here we use a new reduction of the $H$ band data produced by the IRDAP pipeline \citep{2020A&A...633A..64V}, which can separate stellar and instrumental polarization. The polarized signal is consistent with the previous image in \citet{Avenhaus_2018}. The degree of linear polarization of the central and unresolved signal in V4046 Sgr is only 0.13\%, with a systematic uncertainty of 0.05\% due to time-varying atmospheric conditions during the exposures. The angle of polarization is aligned with the disk major disk, as expected given that the target has Av=0.0 \citep{2016ApJ...828...69M} and the entire polarization is dominated by circum rather than inter-stellar material.

@@

Interestingly, the ALMA image also detects faint 1.3\,mm continuum emission near the stellar positions (See the inset in Fig.~\ref{fig:images_vs_simulated}). Since this faint central emission is offset from the star, at the origin of coordinates in the figure, it is probably due to large dust grains. This dust structure is at a distance of only $\sim0\farcs02$ from the binary system, and may be part of some inner structure possibly having to do with the small grain ring called Ring6 introduced in the RT model to match the SED (see Sec.\,\ref{sec:model} below). ***
%match the small grain ring called Ring6 introduced in the RT model to match the SED (see Sec.\,\ref{sec:model} below). If so, then given that the binary phase at the time of the observations is PA=347\,$\degr$, so that the secondary lies roughly northwards of the primary, it may be that the northern side of Ring6 is shadowed, resulting in a seemingly lopsided geometry.

@@

The binary phase reported by \citet{dOrazi} in the scattered-light observation is at a position angle (PA) of 254\,$\degr$, east of north. Using this measurement, the binary phase was calculated at the time of the ALMA observation at a PA of 347\,$\degr$. There is no hint of radio decrements in either Rings13 nor in Ring24 that would match shadowing at this PA, could be explained by inefficient disc cooling compared to the speed of the illumination pattern \citep{Casassus2019MNRAS.486L..58C}.

@@

The observed SED was collected from data in the literature \citep{1988iras....7.....H, 1990A&A...234..230H, Jensen_97, 2000A&A...355L..27H, 2001KFNT...17..409K, 2003yCat.2246....0C, 2007PASJ...59S.369M, 2008PASP..120.1128O, 2010A&A...514A...1I, 2012yCat.2311....0C}, online in \textsc{vizier}, and also using spectrometry data from an archival \textit{Spitzer} IRS spectrum (See Fig.\ref{fig:SED}). The SED will provide us with missing information about the dust population.

@@

For the representation of the stars, we used two Kurucz photospheres models \citep{1979ApJS...40....1K, 1997A&A...318..841C} with T$_{\mathrm{eff},1} =$ 4350\,K, R$_{*,1} =$ 1.064\,R$_{\sun}$, M$_{*,1} =$ 0.90\,M$_{\sun}$ and T$_{\mathrm{eff},2} =$ 4060\,K, R$_{*,2} =$ 1.033\,R$_{\sun}$, M$_{*,2} =$ 0.85\,M$_{\sun}$ respectively and with an accretion rate of log$\,\dot{\mathrm{M}} = -$9.3\,Myr$^{-1}$ for both cases \citep{10.1111/j.1365-2966.2011.19366.x}. The stars were placed at a mutual separation of 0.041\,au, so that their centre of mass coincides with the origin, and oriented at a PA of 250\,$\degr$, so that the secondary lies to the NWW from the primary, thus casting the same shadow as observed by \citet{dOrazi}.

@@

For the vertical structure, \citet{dOrazi} found flaring angles of $\varphi$ = 6.2$\pm$0.6\,$\degr$ for the inner ring and $\varphi$ = 8.5$\pm$1.0\,$\degr$ for the outer one. Our model uses the same values, with two different flaring indexes $\psi$. The separation between the two values were set at $r$ = 19\,au with $\psi=0.2$ for Ring6 and Ring13, and $\psi=0.5$ for Ring24. The scale height is $H_\circ = 0.045$ at $r_0 = 19.5$\,au.

@@

Since the large-dust grains are less coupled to the gas, their distribution has some important differences that require a special parameterisation, with a large inner cavity and scaling factors that account for the larger gap between the rings. For the inner ring, the same profile as the gas is used, but with different values for $R_c$ and $\Sigma_c$. The greatest difference is in the outer ring, where we used a decaying power law modulated by the sum of two Gaussian profiles. In an effort to recreate the break seen in the outer ring between 30.2 and 36.7\,au, we chose a different value for $\gamma$ = -6. For the large dust grains the surface density profile is thus
\begin{multline}
  \Sigma_{\mathrm{large-dust}}(r) = \Sigma_{\mathrm{c}2} \left(\frac{r}{R_{\mathrm{c}2}}\right)^{-\gamma} \\ \, \left\{ \exp\left[-\left(\frac{r}{R_{\mathrm{c}2}}\right)^{2-\gamma}\right] + \mathrm{exp}\left[-\left(\frac{r}{R_{\mathrm{c}2}}\right)^{2}\right]\right\}.
\end{multline}

The inner part of the outer ring follows a different behaviour. Between $R_\mathrm{in}$ and $R_\mathrm{peak}$, the gas density profile is multiplied by an additional factor
\begin{equation}
    \epsilon(r) = \left(\frac{ r - R_\mathrm{in}}{R_\mathrm{peak} - R_\mathrm{in}}\right)^3,
\end{equation}
where $R_\mathrm{in}$ and $R_\mathrm{peak}$ would depend on the dust population, marking the beginning and the maximum density peak of the outer ring. This parameter allow us to model a smoother inner wall of the outer ring.

The two different populations of dust grains corresponds to the small grains, with radii ranging from 0.4 to 1.5\,$\micron$, and the large dust grains range, with radii from 0.4\,$\micron$ to 10\,mm. We computed the dust opacities using the {\tt bhmie} code provided in the \textsc{radmc3d} package, with a mix composed of 70\% silicate and 30\% graphite.

The observed near-far asymmetry in the DPI image is suggestive of a strongly forward-scattering phase function. In order to reproduce a similar asymmetry, we used much larger grains that typically used in the RT modelling of such near-IR data \citep[e.g.,][]{2018MNRAS.477.5104C}.

The simulated DPI image at 1.65\,$\micron$ in Fig.\,\ref{fig:images_vs_simulated} was obtained with the scattering matrix calculated by the {\tt makeopac.py} script provided in the \textsc{radmc3d} package. As a way of reproducing the \textit{H}-band image, we used a different grain size distribution, where we used a Gaussian distribution centred at 0.4\,$\micron$, and smeared out by 30\% in both directions, and using 20 grain size bins within that range. This distribution applies only to generate the NIR image.

@@

The gas and small dust surface density profiles begin at 0.3\,au, out of the zone expected to be cleared by dynamical interactions with the central binary \citep{Art_Lu}, which is expected to be $\sim$0.1\,au. For the small dust grains, we propose a three ringed structure. The innermost ring of the model is located from 5 to 8\,au, this ring, introduced as Ring6 in Sec.\,\ref{sec:Observations}, is not detectable in the IRDIS observations but was required to account for the SED. Then follow the two rings observed by IRDIS: Ring13, from 12 to 19.5\,au, which is surrounded by very narrow 1\,au-wide gap, and then Ring24, that starts 20.5\,au and has its maximum density peak at 25\,au. The small-dust density then decays exponentially with radius.

@@

The simulated 1.3\,mm continuum image reproduces the two observed rings: the faint Ring13 and the brighter Ring24. As the radial profiles obtained from the simulated images of the model closely resemble those deduced from the observations (Fig.~\ref{fig:radprofiles}), we can assume that the model provides an approximation of the disc structure, including the dimensions of Ring13. Accordingly, taking the parametric model values for Ring13 we have that it has a radius of 13.6\,au, a width of 3.7\,au, and a scale height FWHM of the millimetre-size dust ranging from 0.28 to 0.40\,au. The total dust mass of Ring13 is about 3\,M$_{\earth}$ in the model. In turn, in our model Ring24 reaches peak intensity at $\sim$30\,au, then breaks at $\sim$35\,au, as in the observations, and has a total dust mass of $\sim$43\,M$_{\earth}$. The model predictions for the millimetre-size dust population in Ring13 are close to the measurements, with only a 0.14\,au difference in the centroid location of the ring, and a 0.17\,au difference in the width estimation.

As \citet{2018ApJ...869L..46D} explain the stability of a narrow dust ring requires that its height is smaller than its radial extent. Using the parametric model scale height of 0.28\,au at 13.46\,au and the measured width of Ring13 (3.86\,au FWHM, see Section~\ref{sec:Observations}), we have that its width is $\sim$12 times its scale height.

@@

Similarly to \citet{2018ApJ...869L..46D}, the ratio between dust and gas ring width can be linked to important properties:
\begin{equation}\label{eq:ringwidth}
        \frac{w_\mathrm{d}}{w_\mathrm{g}} = \sqrt{\frac{\alpha}{\alpha+\mathrm{St}}}\,,
\end{equation}
where we implicitly assumed the turbulent diffusivity to be equal to the turbulent viscosity. Here, $\alpha$ represents the typical parameter for turbulent viscosity, and $\mathrm{St}$ is the dimensionless Stokes number. Assuming that the most relevant grain size for observed fluxes corresponds to $a = \lambda \mathrm{obs}/(2 \pi)$, and estimating the gas surface density within Ring13 from the RT model to be roughly $20\,\mathrm{g}/\mathrm{cm}^2$, yields $\mathrm{St}\approx 0.03$. With these considerations, equation (\ref{eq:ringwidth}) provides a value of $\alpha \approx 0.01$ for the level of turbulent viscosity. We emphasise that this relies on rough estimates and is additionally restricted to the area within Ring13.

@@

New ALMA 1.3\,mm continuum imaging of the circumbinary disc around V4046\,Sgr were analysed in the context of the available IR imaging and the observed spectral energy distribution using a RT model. The key conclusions are as follows.
\begin{enumerate}
  \item We report the detection of a narrow ring in the 1.3\,mm continuum, with a radius 13.46$\pm$0.43\,au and an estimated width of 3.86$\pm$1.00\,au. The location of this ring is coincident with the inner ring observed in the scattered-light image, revealing that the ring includes around 3\,M$_{\earth}$ millimetre-sized grains. Using the parametric model scale height value ($h= $ 0.28\,au at 13.46\,au) we have that the ring width is roughly 12 times its estimated height.

>>>>

% Abstract of the paper
\begin{abstract}
  The nearby V4046\,Sgr spectroscopic binary hosts a gas-rich disc known for its wide cavity and dusty ring. We present new high resolution ($\sim$50\,mas) ALMA observations of the 1.3\,mm continuum. The comparison of these observations, combined with SPHERE-IRDIS polarized images and a well-sampled spectral energy distribution (SED), against radiative transfer (RT) predictions carried out with the \textsc{radmc3d} package, allow us to propose a physical model for the source. The new ALMA data reveal a very thin ring at a radius of 13.46$\pm$0.43\,au (Ring13), with a marginally resolved radial width of 3.86$\pm$1.00\,au. Ring13 is surrounded by a $\sim$9\,au-wide gap, and it is flanked by a very bright mm outer ring (Ring24) with a sharp inner edge at 24\,au. The steeply decreasing brightness of Ring24 breaks at $\sim$35\,au into a shallow tail. The RT model requires an inner ring at $\sim$6\,au (Ring6) in small dust grains, hiding under the IRDIS coronagraph, and that surrounds an inner circumbinary disc. Faint mm-continuum close ($\sim0\farcs02$) to the stellar position is picked up by ALMA. ***
  %coincident with Ring6 is picked up by ALMA, whose morphology suggests that Ring6 is lopsided or shadowed by the secondary. 
  The previously reported scattered-light shadow of the secondary star is also reproduced by the RT model. The surprisingly thin Ring13 is nonetheless roughly 12 times wider than the model scale height, and could thus be long-lived. The strong near-far disc asymmetry at 1.65\,$\micron$ points at a very forward-scattering phase function, and requires grain radii of no less than 0.4\,$\micron$. 
\end{abstract}

@@

Recent observations of young circumstellar discs have transformed current knowledge of planet formation. The focus of resolved imaging with the Atacama Large Millimetre/submillimetre Array (ALMA) or with 
%VLT/SPHERE 
the current generation of high-contrast images 
has mainly been towards the brighter sources \citep{Andrews2020arXiv200105007A}. The discs around T Tauri Stars (DARTTS) program was first presented in \citet{Avenhaus_2018} with the aim to image eight stars with the Spectro-Polarimeter High-contrast Exoplanet REsearch (SPHERE) \citep{2019A&A...631A.155B}. The sample is not biased towards exceptionally bright and large discs, and consists of only solar-type stars. A second part of the survey increased the number of sources by presenting 21 new images of circumstellar discs \citep{Garufi2020}. The observations revealed diverse structures and morphologies in the scattering surface of these discs. This letter on V4046 Sagittarii (Sgr) is part of the DARTTS survey with ALMA (DARTTS-A) which will present millimetre observations of nine protoplanetary discs previously imaged in polarized scattered light in DARTTS-S.

V4046\,Sgr is a double-lined spectroscopic binary of K-type stars (K5 and K7) with very similar masses of 0.90$\pm$0.05\,M$_{\sun}$ and 0.85$\pm$0.04\,M$_{\sun}$ \citep{Rosenfeld_2012}, on a close ($a \approx 0.041$\,au), circular ($e\leq0.01$) orbit, with an orbital period of 2.42 days \citep{refId0}. It is a member of the $\beta$ Pictoris moving group \citep{Zuckerman_2004}, with an estimated age of 23$\pm$3\,Myr \citep{Mamajek_2014}, and its distance is 72.41$\pm$0.34\,pc \citep{Gaia}. V4046\,Sgr hosts a massive ($\sim$0.1\,M$_{\sun}$) circumbinary disc extending to $\sim$300\,au \citep{Rosenfeld_2013, Rodriguez_2010}, with a rich observable chemical diversity \citep{Kastner_2018}.

@@

V4046 Sgr was observed in differential polarization imaging (DPI) mode with SPHERE/IRDIS on March 13, 2016. All the details of this observation can be found in the first DARTTS-S paper \citep{Avenhaus_2018} where the image was first published. Here we use a new reduction of the $H$ band data produced by the IRDAP pipeline \citep{2020A&A...633A..64V}, which can separate stellar and instrumental polarization. The polarized signal is consistent with the previous image in \citet{Avenhaus_2018}. The degree of linear polarization of the central and unresolved signal in V4046 Sgr is only 0.13\%, with a systematic uncertainty of 0.05\% due to time-varying atmospheric conditions during the exposures. The angle of polarization is aligned with the disk major disk, as expected given that the target has Av=0.0 \citep{2016ApJ...828...69M} and the entire polarization is dominated by circumstellar rather than inter-stellar material.

@@

Interestingly, the ALMA image also detects faint 1.3\,mm continuum emission near the stellar positions (See the inset in Fig.~\ref{fig:images_vs_simulated}). Since this faint central emission is offset from the star, at the origin of coordinates in the figure, it is probably due to large dust grains. This dust structure is at a distance of only $\sim0\farcs02$ from the binary system, and may be part of some inner structure possibly having to do with the small grain ring called Ring6 introduced in the RT model to match the SED (see Sec.~\ref{sec:model} below). ***
%match the small grain ring called Ring6 introduced in the RT model to match the SED (see Sec.\,\ref{sec:model} below). If so, then given that the binary phase at the time of the observations is PA=347\,$\degr$, so that the secondary lies roughly northwards of the primary, it may be that the northern side of Ring6 is shadowed, resulting in a seemingly lopsided geometry.

@@

The binary phase reported by \citet{dOrazi} in the scattered-light observation is at a position angle (PA) of 254\,$\degr$, east of north. Using this measurement, the binary phase was calculated at the time of the ALMA observation at a PA of 347\,$\degr$. There is no hint of radio decrements in either Rings13 nor in Ring24 that would match shadowing at this PA. This could be explained by inefficient disc cooling compared to the speed of the illumination pattern \citep{Casassus2019MNRAS.486L..58C}.

@@

The observed SED was collected from data in the literature \citep{1988iras....7.....H, 1990A&A...234..230H, Jensen_97, 2000A&A...355L..27H, 2001KFNT...17..409K, 2003yCat.2246....0C, 2007PASJ...59S.369M, 2008PASP..120.1128O, 2010A&A...514A...1I, 2012yCat.2311....0C}, online in \textsc{vizier}, and also using spectrometry data from an archival \textit{Spitzer} IRS spectrum (See Fig.~\ref{fig:SED}). The SED will provide us with missing information about the dust population.

@@

For the representation of the stars, we used two Kurucz photospheres models \citep{1979ApJS...40....1K, 1997A&A...318..841C} with T$_{\mathrm{eff},1} =$ 4350\,K, R$_{*,1} =$ 1.064\,R$_{\sun}$, M$_{*,1} =$ 0.90\,M$_{\sun}$ and T$_{\mathrm{eff},2} =$ 4060\,K, R$_{*,2} =$ 1.033\,R$_{\sun}$, M$_{*,2} =$ 0.85\,M$_{\sun}$ respectively and with an accretion rate of $\mathrm{log}(\,\dot{\mathrm{M}/(\mathrm{M}_{\sun}\,\mathrm{yr^{-1}})}) = -$9.3 for both cases \citep{10.1111/j.1365-2966.2011.19366.x}. The stars were placed at a mutual separation of 0.041\,au, so that their centre of mass coincides with the origin, and oriented at a PA of 250\,$\degr$, so that the secondary lies to the NWW from the primary, thus casting the same shadow as observed by \citet{dOrazi}.

@@

For the vertical structure, \citet{dOrazi} found flaring angles of $\varphi$ = 6.2$\pm$0.6\,$\degr$ for the inner ring and $\varphi$ = 8.5$\pm$1.0\,$\degr$ for the outer one. Our model uses the same values, with two different flaring indexes $\psi$. The separation between the two values were set at $r$ = 19\,au with $\psi=0.2$ for Ring6 and Ring13, and $\psi=0.5$ for Ring24. The scale height is $H_\circ = 0.045$\,au at $r_\circ = 19.5$\,au.

@@

Since the large dust grains are less coupled to the gas, their distribution has some important differences that require a special parameterisation, with a large inner cavity and scaling factors that account for the larger gap between the rings. For the inner ring, the same profile as the gas is used, but with different values for $R_c$ and $\Sigma_c$. The greatest difference is in the outer ring, where we used a decaying power law modulated by the sum of two Gaussian profiles. In an effort to recreate the break seen in the outer ring between 30.2 and 36.7\,au, we chose a different value for $\gamma$ = -6. For the large dust grains the surface density profile is thus
\begin{multline}
  \Sigma_{\mathrm{large-dust}}(r) = \Sigma_{\mathrm{c}2} \left(\frac{r}{R_{\mathrm{c}2}}\right)^{-\gamma} \\ \, \left\{ \exp\left[-\left(\frac{r}{R_{\mathrm{c}2}}\right)^{2-\gamma}\right] + \mathrm{exp}\left[-\left(\frac{r}{R_{\mathrm{c}2}}\right)^{2}\right]\right\}.
\end{multline}

The inner part of the outer ring follows a different behaviour. Between $R_\mathrm{in}$ and $R_\mathrm{peak}$, marking the beginning and the maximum density peak of the outer ring respectively, the gas density profile is multiplied by an additional factor
\begin{equation}
    \epsilon(r) = \left(\frac{ r - R_\mathrm{in}}{R_\mathrm{peak} - R_\mathrm{in}}\right)^3.
\end{equation}
$R_\mathrm{in}$ and $R_\mathrm{peak}$ would depend on the dust population, with $R_\mathrm{in}$ = 20.5\,au and $R_\mathrm{peak}$ = 25\,au for the small dust grains and $R_\mathrm{in}$ = 22\,au and $R_\mathrm{peak}$ = 27.7\,au for the large dust grains. This parameter allow us to model a smoother inner wall of the outer ring.

The two different populations of dust grains correspond to small grains, with radii ranging from 0.4 to 1.5\,$\micron$, and large dust grains, with radii from 0.4\,$\micron$ to 10\,mm. We computed the dust opacities using the {\tt bhmie} code provided in the \textsc{radmc3d} package, with a mix composed of 70\% silicate and 30\% graphite.

The observed near-far asymmetry in the DPI image is suggestive of a strongly forward-scattering phase function. In order to reproduce a similar asymmetry, we used much larger grains than typically used in the RT modelling of such near-IR data \citep[e.g.,][]{2018MNRAS.477.5104C}.

The simulated DPI image at 1.65\,$\micron$ in Fig.~\ref{fig:images_vs_simulated} was obtained with the scattering matrix calculated by the {\tt makeopac.py} script provided in the \textsc{radmc3d} package. As a way of reproducing the \textit{H}-band image, we used a different grain size distribution, where we used a Gaussian distribution centred at 0.4\,$\micron$, and smeared out by 30\% in both directions, and using 20 grain size bins within that range. This distribution applies only to generate the NIR image.

@@

The gas and small dust surface density profiles begin at 0.3\,au, out of the zone expected to be cleared by dynamical interactions with the central binary \citep{Art_Lu}, which is expected to be $\sim$0.1\,au. For the small dust grains, we propose a three ringed structure. The innermost ring of the model is located from 5 to 8\,au, this ring, introduced as Ring6 in Sec.~\ref{sec:Observations}, is not detectable in the IRDIS observations but was required to account for the SED. Then follow the two rings observed by IRDIS: Ring13, from 12 to 19.5\,au, which is surrounded by very narrow 1\,au-wide gap, and then Ring24, that starts 20.5\,au and has its maximum density peak at 25\,au. The small-dust density then decays exponentially with radius.

@@

The simulated 1.3\,mm continuum image reproduces the two observed rings: the faint Ring13 and the brighter Ring24. As the radial profiles obtained from the simulated images of the model closely resemble those deduced from the observations (Fig.~\ref{fig:radprofiles}), we can assume that the model provides a good approximation of the disc structure, including the dimensions of Ring13. Accordingly, taking the parametric model values for Ring13 we have that it has a radius of 13.6\,au, a width of 3.7\,au, and a scale height FWHM of the millimetre-size dust ranging from 0.28 to 0.40\,au. The total dust mass of Ring13 is about 3\,M$_{\earth}$ in the model. In turn, in our model Ring24 reaches peak intensity at $\sim$30\,au, then breaks at $\sim$35\,au, as in the observations, and has a total dust mass of $\sim$43\,M$_{\earth}$. The model predictions for the millimetre-size dust population in Ring13 are close to the measurements, with only a 0.14\,au difference in the centroid location of the ring, and a 0.17\,au difference in the width estimation.
%As \citet{2018ApJ...869L..46D} explain the stability of a narrow dust ring requires that its height is smaller than its radial extent. 
Using the parametric model scale height of 0.33\,au at 13.46\,au and the measured width of Ring13 (3.86\,au FWHM, see Section~\ref{sec:Observations}), we have that its width is $\sim$12 times its scale height.

@@

Similarly to \citet{2018ApJ...869L..46D}, the ratio between dust and gas ring width can be linked to important properties:
\begin{equation}\label{eq:ringwidth}
        \frac{w_\mathrm{d}}{w_\mathrm{g}} = \sqrt{\frac{\alpha}{\alpha+\mathrm{St}}}\,,
\end{equation}
where we implicitly assumed the turbulent diffusivity to be equal to the turbulent viscosity. Here, $\alpha$ represents the typical parameter for turbulent viscosity, and $\mathrm{St}$ is the dimensionless Stokes number. Assuming that the most relevant grain size for observed fluxes corresponds to $a = \lambda_{\mathrm{obs}}/(2 \pi)$, and estimating the gas surface density within Ring13 from the RT model to be roughly $20\,\mathrm{g}/\mathrm{cm}^2$, yields $\mathrm{St}\approx 0.03$. With these considerations, equation~\ref{eq:ringwidth} provides a value of $\alpha \approx 0.01$ for the level of turbulent viscosity. We emphasise that this relies on rough estimates and is additionally restricted to the area within Ring13.

@@

New ALMA 1.3\,mm continuum imaging of the circumbinary disc around V4046\,Sgr were analysed in the context of the available IR imaging and the observed spectral energy distribution using a RT model. The key conclusions are as follows.
\begin{enumerate}
  \item We report the detection of a narrow ring in the 1.3\,mm continuum, with a radius 13.46$\pm$0.43\,au and an estimated width of 3.86$\pm$1.00\,au. The location of this ring is coincident with the inner ring observed in the scattered-light image, revealing that the ring includes around 3\,M$_{\earth}$ millimetre-sized grains. Using the parametric model scale height value ($H= $ 0.33\,au at 13.46\,au) we have that the ring width is roughly 12 times its estimated height.
